---
title: "A Novel Hybrid Approach to Contraceptive Demand Forecasting: Integrating Point Predictions with Probabilistic Distributions"

author:
  - name: Harsha Chamara Hewage
    email: HalgamuweHewageHR@cardiff.ac.uk
    affiliations:
        - id: dl4sg
          name: Cardiff University
          department: Data Lab for Social Good Group, Cardiff Business School
          country: United Kingdom
          postal-code: CF10 3EU
    attributes:
        corresponding: true
        
  - name: Bahman Rostami-Tabar
    email: rostami-tabarb@cardiff.ac.uk
    affiliations:
        - id: dl4sg
          name: Cardiff University
          department: Data Lab for Social Good Group, Cardiff Business School
          country: United Kingdom
          postal-code: CF10 3EU
    
  - name: Aris Syntetos
    email: syntetosa@cardiff.ac.uk
    affiliations:
        - id: carbs
          name: Cardiff University
          department: Cardiff Business School
          country: United Kingdom
          postal-code: CF10 3EU
          
  - name: Federico Liberatore
    email: liberatoref@cardiff.ac.uk
    affiliations:
        - id: comsc
          name: Cardiff University
          department: Cardiff School of Computer Science & Informatics
          country: United Kingdom
          postal-code: CF24 4AG
          
  - name: Glenn Milano
    email: gmilano@usaid.gov
    affiliations:
        - id: usaid
          name: United States Agency for International Development
          department: Bureau for Global Health
          country: United States
          postal-code: 20523
          

abstract: | 
  Accurate demand forecasting is vital for ensuring reliable access to contraceptive products, supporting key processes like procurement, inventory, and distribution. However, forecasting contraceptive demand in developing countries presents challenges, including incomplete data, poor data quality, and the need to account for multiple geographical and product factors. Current methods often rely on simple forecasting techniques, which fail to capture demand uncertainties arising from these factors, warranting expert involvement. Our study aims to improve contraceptive demand forecasting by combining probabilistic forecasting methods with expert knowledge. We developed a hybrid model that combines point forecasts from domain-specific model with probabilistic distributions from statistical and machine learning approaches, enabling human input to fine-tune and enhance the system-generated forecasts. This approach helps address the uncertainties in demand and is particularly useful in resource-limited settings. We evaluate different forecasting methods, including time series, Bayesian, machine learning, and foundational time series methods alongside our new hybrid approach. By comparing these methods, we provide insights into their strengths, weaknesses, and computational requirements. Our research fills a gap in forecasting contraceptive demand and offers a practical framework that combines algorithmic and human expertise. Our proposed model can also be generalized to other humanitarian contexts with similar data patterns.
  
keywords:
  - family planning supply chain
  - hybrid forecasting
  - forecast distributions
  - contraceptive demand
  - forecast combination
  
bibliography: bibliography.bib

execute:
  echo: false
  message: false
  warning: false
  cache: true
pdf-engine: pdflatex
prefer-html: true
format:
  html:
    number-sections: true
  docx:
    number-sections: true
  tandf-pdf:
    keep-tex: true
    include-in-header:
      text: |
        \usepackage{todonotes,mathtools,bm,amsmath,mathpazo}
        \mathtoolsset{showonlyrefs}
        \setlength{\parindent}{0cm}
    journal:
      #name: Journal of Service Research
      #formatting: preprint
      model: 3p
      cite-style: authoryear
---


```{r}
#| label: setup
#| context: setup
#| include: false

library(tidyverse)
library(fpp3)
library(knitr)
library(kableExtra)
library(ggthemes)
library(patchwork)
library(scales)
library(ggdist)
library(tsutils)


# Color blind color palettes
options(
  ggplot2.discrete.colour = c("#D55E00", "#0072B2", "#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442"),
  ggplot2.discrete.fill = c("#D55E00", "#0072B2", "#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442")
)
# Few theme
ggplot2::theme_set(ggthemes::theme_few())

```

```{r intro, child="01-introduction.qmd"}
```

```{r literature, child="02-literature.qmd"}
```

```{r literature, child="03-proposed_model.qmd"}
```

```{r experiment, child="04-experiment.qmd"}
```

```{r results, child="05-results.qmd"}
```

```{r discussion, child="06-discussion.qmd"}
```

```{r conclusion, child="07-conclusion.qmd"}
```


# Data availability statement {#reproducibility .unnumbered}

The R and Python code, along with the data required to reproduce all results in this paper, will be made available in a GitHub repository upon acceptance of the paper.

# Acknowledgements {#acknowledgements .unnumbered}

We would like to thank Laila Akhlaghi from John Snow, Inc. for her advice and feedback.

# Disclosure statement {.unnumbered}

The authors report there are no competing interests to declare.

# References {#references .unnumbered}

::: {#refs}
:::

\pagebreak

# Appendix {#appendix .unnumbered}

## Appendix 1 {#appendix1 .appendix .unnumbered}

```{r appendix1, child="08-appendix_fig.qmd"}
```

\pagebreak 

## Appendix 2 {#appendix2 .appendix .unnumbered}

```{r appendix2, child="08-appendix_tbl.qmd"}
```

